---
author: Qiran Chen
title: "p8105_hw2_qc2336"
output: github_document
date: "2023-09-26"
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Problem 1
```{r cars}
pol_month = read_csv(file = "./data/pols-month.csv")|>
  janitor::clean_names()|>
  separate(mon, into = c("year", "month","day"), convert = TRUE) |>
  mutate(year=as.integer(year),
  month=month.abb[as.numeric(month)],
  day=as.integer(day))|>
  mutate(president=recode(prez_dem, `1`="dem",`0`="gop"))|>
  select(-prez_dem,-prez_gop, -day)
  
pol_month
```


```{r}
snp = read_csv(file = "./data/snp.csv")|>
  janitor::clean_names()|>
  separate(date,into = c("month","day","year"),convert=TRUE)|>
  mutate(month=month.abb[as.integer(month)],
         year=as.numeric(year),
        year=ifelse(year<=20,year+2000,year+1900))|>
  arrange(year,month)|>
  select(year,month,close,-c("day"))
  

snp

```

```{r}
unemployment=read_csv(file = "./data/unemployment.csv")|>
  pivot_longer(Jan:Dec,
               names_to = 'month',
               values_to = 'per_unemploy')|>
  rename(year=Year)

unemployment
  
```

```{r}
pol_snp=left_join(pol_month,snp,by=c("year","month"))

pol_snp_unemp=left_join(pol_snp,unemployment,by=c("year","month"))

pol_snp_unemp
```

>> The pol_month dtata set contains the number of republican and democrates in senate, governor and presidential positions in 1947-2015, it has 822 observations and 9 variables including whether the president is dem or rep on the date. The snp data shows the closing value of stock in 1950-2015, it has 787 observatins and 10 variables, the most important variable is the closing value. And the umemployment data shows the unempolyment rate in 1948-2015, the variable per_unemploy shows the unemploymet percentage of each date. The pol_snp_unemp data frame has 822 observation of 11 variables with the variables in the aforementioned dataframes combined from year 1947-2015.

## Problem 2


```{r pressure, echo=FALSE}
Mr_trashwheel = readxl::read_excel("./data/202207 Trash Wheel Collection Data.xlsx", sheet="Mr. Trash Wheel",range = "A2:N550")|>
  janitor::clean_names() |>
  mutate(homes_powered=(as.numeric(weight_tons)*500/30),
         year=as.numeric(year),
         type="Mr_Trashwheel")|>
  drop_na(dumpster)
  
Mr_trashwheel
```

```{r}
Pf_trashwheel = readxl::read_excel("./data/202207 Trash Wheel Collection Data.xlsx", sheet="Professor Trash Wheel" ,range ="A2:M97")|>
  janitor::clean_names() |>
  drop_na(dumpster)|>
  mutate(type="Pf_trashwheel")

Pf_trashwheel
```

```{r}
Gwynnda = readxl::read_excel("./data/202207 Trash Wheel Collection Data.xlsx", sheet="Gwynnda Trash Wheel" ,range ="A2:K110")|>
  janitor::clean_names() |>
  drop_na(dumpster)|>
  mutate(type="gwynnda")

Gwynnda
```
```{r}

trash_tidy = 
  bind_rows(Pf_trashwheel, Gwynnda, Mr_trashwheel) |>
  select(type,dumpster,everything())

trash_tidy

```
>> The three data has the relative similar variables, the key variables are the weight tons of the trash collected and the homes powers. The Mr. Trashwheel data set has 547 observation and 15 variable, Gwynnda has 12 variable and 106 observations, Pr_trashwheel has 94 observations and 14 variables. The difference is on the kind of trash collected, variables such as plastic bags are not measured in the other datasets. The total trash collected by pf_trashwheel is 190.12 and total for gwynnda is 310.78

```{r}
sum(Pf_trashwheel$weight_tons)
sum(Gwynnda$weight_tons)
```

## Problem 3
```{r}
baseline = read_csv(file = "./data/data_mci/MCI_baseline.csv", skip=1)|>
  janitor::clean_names()|>
  mutate(sex=factor(sex,levels=c(0,1),labels=c("Female","Male")),
         apoe4=factor(apoe4,levels=c(0,1),labels=c("Non-carrier","Carrier")),
         age_at_onset=as.numeric(age_at_onset))|>
  drop_na(age_at_onset)
  

baseline
```
>> The data set has six variabled. Current age, age at onset, sex, apoe4 gene, and id. In this data set, the important steps are to change the binary data sex and apoe 4 into chracters. And change the age of onset into numeric data so it can drop the missing value.

```{r}
baseline|>
  count(sex, apoe4)
```
>> 483 people are recruited but only `count(baseline)` developed mci. The mean baseline age is 70.26. `30/40` of the women are APOE4 carriers. 

```{r}
amyloid = read_csv(file = "./data/data_mci/mci_amyloid.csv",skip=1)|>
  janitor::clean_names()|>
  mutate(baseline=as.numeric(baseline),
         time_2=as.numeric(time_2),
         time_4=as.numeric(time_4),
         time_6=as.numeric(time_6),
         time_8=as.numeric(time_8))|>
  rename(id=study_id)


amyloid
```

>> This data has six variable, id, baseline, time2,4,6,8. It's mostly numerical but it shows it is character in data set. During the tidying this is changed by using mutate. The study id is renamed as id so it is easier for the table to merge.

```{r}
base_amyloid<-left_join(baseline,amyloid,by="id")

write_csv(base_amyloid, 'base_amyloid.csv')

base_amyloid
```
>> Some participants are removed in the baseline dataset becomes they never developed mci. When left joined the dataset has 97 observation of subjects that developed mci during the study and their longtitudinal data combined with data in baseline. It has 11 variables that is show in the two table.
