---
title: "p8105_hw2_qc2336"
output: html_document
date: "2023-09-26"
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Problem 1
```{r cars}
pol_month = read_csv(file = "./data/pols-month.csv")|>
  janitor::clean_names()|>
  separate(mon, into = c("year", "month","day"), convert = TRUE) |>
  mutate(year=as.integer(year),
  month=month.abb[as.numeric(month)],
  day=as.integer(day))|>
  mutate(president=recode(prez_dem, `1`="dem",`0`="gop"))|>
  select(-prez_dem,-prez_gop, -day)
  
pol_month
```


```{r}
snp = read_csv(file = "./data/snp.csv")|>
  janitor::clean_names()|>
  separate(date,into = c("month","day","year"),convert=TRUE)|>
  mutate(month=month.abb[as.integer(month)],
         year=as.numeric(year),
        year=ifelse(year<=20,year+2000,year+1900))|>
  arrange(year,month)|>
  select(year,month,close,-c("day"))
  

snp

```

```{r}
unemployment=read_csv(file = "./data/unemployment.csv")|>
  pivot_longer(Jan:Dec,
               names_to = 'month',
               values_to = 'per_unemploy')|>
  rename(year=Year)

unemployment
  
```

```{r}
pol_snp=left_join(pol_month,snp,by=c("year","month"))

pol_snp_unemp=left_join(pol_snp,unemployment,by=c("year","month"))

pol_snp_unemp
```

>> The pol_month dtata set contains the number of republican and democrates in senate, governor and presidential positions in 1947-2015, it has 822 observations and 9 variables including whether the president is dem or rep on the date. The snp data shows the closing value of stock in 1950-2015, it has 787 observatins and 10 variables, the most important variable is the closing value. And the umemployment data shows the unempolyment rate in 1948-2015, the variable per_unemploy shows the unemploymet percentage of each date. The pol_snp_unemp data frame has 822 observation of 11 variables with the variables in the aforementioned dataframes combined from year 1947-2015.

## Problem 2


```{r pressure, echo=FALSE}
Mr_trashwheel = readxl::read_excel("./data/202207 Trash Wheel Collection Data.xlsx", sheet="Mr. Trash Wheel",range = "A2:N550")|>
  janitor::clean_names() |>
  mutate(homes_powered=(as.numeric(weight_tons)*500/30))|>
  drop_na(dumpster)
  
Mr_trashwheel

Pr_trashwheel = readxl::read_excel("./data/202207 Trash Wheel Collection Data.xlsx", sheet="Professor Trash Wheel" ,range ="A2:M97")|>
  janitor::clean_names() |>
  drop_na(dumpster)

```
```{r}

```

